\documentclass[12pt]{article}
\usepackage{geometry}
\geometry{a4paper, margin=1in}
\usepackage{amsmath, amssymb, graphicx, xcolor}
\usepackage{enumitem}
\usepackage{hyperref}

\title{THESIS PROPOSAL}
\author{Your Name}
\date{Month, Year}

\begin{document}

\maketitle

\section*{Title}
Title of the Thesis Proposal

\section*{Abstract}
\textbf{Background of Research:}

Briefly describe the background of the research.

\textbf{Objective of Research:}

State the main objective(s) of the research.

\textbf{Methods:}

Outline the methods to be used in the study.

\textbf{Expected Results:}

Summarize the expected findings or contributions of the study.

\section{Introduction}
\begin{itemize}
  \item \textbf{Background:} Provide context and background for the research topic.
  \item \textbf{State of the Art:} Review the current state of the art in the relevant field.
  \item \textbf{Gap Analysis:} Identify gaps in existing research.
  \item \textbf{Research Question:} Clearly state the research question(s).
  \item \textbf{Research Objective:} Define the objectives of the study.
  \item \textbf{Limitation:} Describe the scope and limitations of the research.
  \item \textbf{Hypothesis:} State the research hypothesis.
\end{itemize}

\section{Literature Review}
Discuss a minimum of 20 citations related to the research topic. Compare existing methods, models, and approaches. Identify research gaps and justify the proposed work.

\section{Methodology}
\begin{itemize}
  \item \textbf{Research Design:} Outline the overall design of the study.
  \item \textbf{Data Collection:} Describe the data sources and collection methods.
  \item \textbf{Data Preprocessing:} Explain data cleaning and preprocessing techniques.
  \item \textbf{Data Processing:} Discuss data processing steps and tools.
  \item \textbf{Data Analysis:} Specify data analysis techniques.
  \item \textbf{Data Visualization:} Outline data visualization methods.
  \item \textbf{Data Validity:} Discuss validity and reliability considerations.
\end{itemize}

\end{document}
